Linear instability underlying synchronous waving of marine grass in flow

### 2014-08-30 Paper/grass.tex
Combined fig showing critical Re and k

We use a mean field model for the coupling between the flow and the canopy. Drag force exerted by the grass is approximated via a 
continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance\cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06} as
\begin{equation}
\rho \left(\bu_{t}+\bu.\grad\bu \right) = -\grad P+\mu\grad^{2}\bu +\mathbf{f}+\rho\mathbf{g}, \quad \grad\cdot\bu = 0,
\label{navier-stokes}
\end{equation}
%\begin{equation}
% \mathbf{f}=-N\mathbf{f_{d}}
%\end{equation}
where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, $N_g$ the grass number density per unit area, $\rho$ the fluid density, $\mathbf{u}$ the velocity, 
$P$ the pressure, $\mu$ the turbulent viscosity and $\mathbf{g}$ the acceleration due to gravity. The Reynolds number of the flow on the scale of a grass blade is expected to 
be O($10^2$-$10^3$), and hence the drag force is dominated by form drag, and is modeled 
as $\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\hat{n}+C_{T}\rho u_{T}^{2}d\hat{t}$ where $d$ is average width of grass blade
$C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively which are set to zero outside the grass; $\bu_{T}$, $\bu_{N}$ are velocity vector along and
normal to grass while $\hat{t},\hat{n}$ being unit vector along and normal to grass. We expect $C_T \ll C_N$ and take $C_T=0$ for rest of analysis. In the field, both $C_N \& N_g$ 
vary with distance from the bottom but we do not expect these variation to be central to the mechanism and therefore take $C_N$ \& $N_g$ to be constants. Based on previous experiments\cite{Vivoni98}, we take $C_N \approx 1$.

@@

Considering small perturbations $u, v, p$ about the base profile $U$ and $P$ respectively, equation \eqref{navier-stokes} expanded to linear order in the perturbations yields
\begin{equation}
\begin{split}
\frac{\del u}{\del t}+U\frac{\del u}{\del x}+v\frac{\del U}{\del y} &= -\frac{1}{\rho}\frac{\del p}{\del x}+\frac{\mu}{\rho}\nabla^2u-2S(y) C_{N}dN_{g}Uu, \\
\frac{\del v}{\del  t}+ U\frac{\del v}{\del x} &= -\frac{1}{\rho}\frac{\del p}{\del y}+\frac{\mu}{\rho}\nabla^2v, \hspace{0.3cm} \nabla\cdot\bu=0.
 %\nabla\cdot \bu &= 0
\end{split}
\end{equation}
These equations can be non-dimensionalized using half channel height $H$, velocity $U_0$, and the associated advection time $H/U_0$ . 
We are left with three non-dimensional parameters in the model, viz. the Reynolds number $\Rey$, the vegetation drag parametrized by $\Ndg$, 
and the submergence ratio of the grass $h_g/H$. We also parametrize the grass density through the boundary layer thickness $\delta/H$ to help elucidate the mechanism of the instability. With these scaling, and using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a wave solution of the form $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma t}$ to  obtain a modified Orr-Sommerfield equation. The eigenvalue $\sigma$ for a given $k$ provides growth rate and frequency associated with a perturbation of wavelength $2\pi/k$ through the solution of
% \small
\begin{equation}
\begin{split}
\left(D^2 -k^{2} \right)^2\phi &= \Rey \left[ \left({\sigma}+ikU\right) \left(D^2-k^2\right) -ikU_{yy}\right]\phi \\
&+D\left(2R \Ndg S(y) U D \phi\right),
\label{Orr-somerfield}
\end{split}
\end{equation}
where $D=d/dy$, and subject to the boundary conditions $D\phi = D^2\phi = 0$ at $y=\pm 1$. 
\begin{figure}[htb]
\includegraphics[]{Critical_Re_vs_delta_noshear}
\caption{Critical Reynolds number with flow for different grass height as a function of inverse shear thickness}
\label{Re_vs_delta}
\end{figure}
A threshold in $\Rey$ for the onset of the instability naturally emerges from the solution of \eqref{Orr-somerfield}. The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on the $\delta/H$ and $\hg/H$ is shown in figure \ref{Re_vs_delta}, and compared with experimental observation of the threshold\cite{Ghisal02} without any fitting parameters.

@@

\begin{figure}[htb]
\includegraphics[]{K_vs_shear_width_noshear}
\caption{Critical Wavenumber for different grass height as function of shear layer thickness}
\label{K_vs_shear}
\end{figure}
\begin{figure*}

%\subfigure[]{\includegraphics[scale = 0.24]{Set3_dens31_imgsc}}
%\subfigure[]{\includegraphics[scale = 0.24]{Set3_dens32_imgsc}}
%\subfigure[]{\includegraphics[scale = 0.24]{Set3_dens34_imgsc}}
%\subfigure[]{\includegraphics[scale = 0.24]{Set3_dens38_imgsc}}

\begin{tabular}{cccc}
{\includegraphics[scale = 0.72]{Set4_dens28_imgsc}} &
{\includegraphics[scale = 0.72]{Set4_dens30_imgsc}} &
{\includegraphics[scale = 0.72]{Set4_dens32_imgsc}} &
{\includegraphics[scale = 0.72]{Set4_dens34_imgsc}} \\
{\includegraphics[scale = 0.72]{Set5_dens38_imgsc}} &
{\includegraphics[scale = 0.72]{Set5_dens40_imgsc}} &
{\includegraphics[scale = 0.72]{Set5_dens42_imgsc}} &
{\includegraphics[scale = 0.72]{Set5_dens46_imgsc}} \\
{\includegraphics[scale = 0.72]{Set6_dens32_imgsc}} &
{\includegraphics[scale = 0.72]{Set6_dens34_imgsc}} &
{\includegraphics[scale = 0.72]{Set6_dens36_imgsc}} &
{\includegraphics[scale = 0.72]{Set6_dens38_imgsc}}
\end{tabular}

@@

On the other hand, Mode 1 localizes with a length scale $\delta$ around the grass tip, and exhibits a different asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto \sqrt{\Ndg}$) at the threshold. This limit can be understood through asymptotically balancing the magnitude of the advection term ($\Rey/\delta^3 \sim \Rey^2 \Ndg $) with that of the viscous term (or the vegetation drag term, which has the same magnitude) in the boundary layer ($\delta^{-4} \sim (\Rey \Ndg)^{4/3}$). The terms balance when $\Rey \sim \sqrt{\Ndg}$, explaining the asymptote. The resulting localization of the the mode structure to a length of $O(\delta)$, is shown in figure \ref{Asymptotic_mode}, supporting this argument.

>>>>

We use a mean field model for the coupling between the flow and the canopy. Drag force exerted by the grass is approximated via a 
continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance\cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06} as
\begin{equation}
\rho \left(\bu_{t}+\bu.\grad\bu \right) = -\grad P+\mu\grad^{2}\bu +\mathbf{f}+\rho\mathbf{g}, \quad \grad\cdot\bu = 0,
\label{navier-stokes}
\end{equation}
where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, $N_g$ the grass number density per unit area, $\rho$ the fluid density, $\mathbf{u}$ the velocity, 
$P$ the pressure, $\mu$ the turbulent viscosity and $\mathbf{g}$ the acceleration due to gravity. The Reynolds number of the flow on the scale of a grass blade is expected to 
be O($10^2$-$10^3$), and hence the drag force is dominated by form drag, and is modeled 
as $\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\hat{n}+C_{T}\rho u_{T}^{2}d\hat{t}$ where $d$ is average width of grass blade
$C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively which are set to zero outside the grass; $\bu_{T}$, $\bu_{N}$ are velocity vector along and
normal to grass while $\hat{t},\hat{n}$ being unit vector along and normal to grass. We expect $C_T \ll C_N$ and take $C_T=0$ for rest of analysis. In the field, both $C_N \& N_g$ 
vary with distance from the bottom but we do not expect these variation to be central to the mechanism and therefore take $C_N$ \& $N_g$ to be constants. Based on previous experiments\cite{Vivoni98}, we take $C_N \approx 1$.

@@

Considering small perturbations $u, v, p$ about the base profile $U$ and $P$ respectively, equation \eqref{navier-stokes} expanded to linear order in the perturbations yields
\begin{equation}
\begin{split}
\frac{\del u}{\del t}+U\frac{\del u}{\del x}+v\frac{\del U}{\del y} &= -\frac{1}{\rho}\frac{\del p}{\del x}+\frac{\mu}{\rho}\nabla^2u-2S(y) C_{N}dN_{g}Uu, \\
\frac{\del v}{\del  t}+ U\frac{\del v}{\del x} &= -\frac{1}{\rho}\frac{\del p}{\del y}+\frac{\mu}{\rho}\nabla^2v, \hspace{0.3cm} \nabla\cdot\bu=0.
 %\nabla\cdot \bu &= 0
\end{split}
\end{equation}
These equations can be non-dimensionalized using half channel height $H$, velocity $U_0$, and the associated advection time $H/U_0$ . 
We are left with three non-dimensional parameters in the model, viz. the Reynolds number $\Rey$, the vegetation drag parametrized by $\Ndg$, 
and the submergence ratio of the grass $h_g/H$. We also parametrize the grass density through the boundary layer thickness $\delta/H$ to help elucidate the mechanism of the instability. With these scaling, and using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a wave solution of the form $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma t}$ to  obtain a modified Orr-Sommerfield equation. The eigenvalue $\sigma$ for a given $k$ provides growth rate and frequency associated with a perturbation of wavelength $2\pi/k$ through the solution of
% \small
\begin{equation}
\begin{split}
\left(D^2 -k^{2} \right)^2\phi &= \Rey \left[ \left({\sigma}+ikU\right) \left(D^2-k^2\right) -ikU_{yy}\right]\phi \\
&+D\left(2R \Ndg S(y) U D \phi\right),
\label{Orr-somerfield}
\end{split}
\end{equation}
where $D=d/dy$, and subject to the boundary conditions $D\phi = D^2\phi = 0$ at $y=\pm 1$.

\begin{figure}[htb]
\begin{center}
\includegraphics[]{Critical_Re_vs_delta_noshear} \\
\vspace{-6mm} \hspace{-3mm}
\includegraphics[]{K_vs_shear_width_noshear}
\end{center}
\caption{Critical Reynolds number with flow for different grass height as a function of inverse shear thickness}
\label{Re_vs_delta}
\end{figure}

A threshold in $\Rey$ for the onset of the instability naturally emerges from the solution of \eqref{Orr-somerfield}. The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on the $\delta/H$ and $\hg/H$ is shown in figure \ref{Re_vs_delta}, and compared with experimental observation of the threshold\cite{Ghisal02} without any fitting parameters.

@@

\begin{figure*}
\begin{tabular}{cccc}
{\includegraphics[scale = 0.72]{Set4_dens28_imgsc}} &
{\includegraphics[scale = 0.72]{Set4_dens30_imgsc}} &
{\includegraphics[scale = 0.72]{Set4_dens32_imgsc}} &
{\includegraphics[scale = 0.72]{Set4_dens34_imgsc}} \\
{\includegraphics[scale = 0.72]{Set5_dens38_imgsc}} &
{\includegraphics[scale = 0.72]{Set5_dens40_imgsc}} &
{\includegraphics[scale = 0.72]{Set5_dens42_imgsc}} &
{\includegraphics[scale = 0.72]{Set5_dens46_imgsc}} \\
{\includegraphics[scale = 0.72]{Set6_dens32_imgsc}} &
{\includegraphics[scale = 0.72]{Set6_dens34_imgsc}} &
{\includegraphics[scale = 0.72]{Set6_dens36_imgsc}} &
{\includegraphics[scale = 0.72]{Set6_dens38_imgsc}}
\end{tabular}

@@

On the other hand, Mode 1 asymptotically localizes with a length scale $\delta$ around the grass tip, and exhibits a different asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto \sqrt{\Ndg}$) at the threshold. This limit can be understood through asymptotically balancing the magnitude of the advection term ($\Rey/\delta^3 \sim \Rey^2 \Ndg $) with that of the viscous term (or the vegetation drag term, which has the same magnitude) in the boundary layer ($\delta^{-4} \sim (\Rey \Ndg)^{4/3}$). The terms balance when $\Rey \sim \sqrt{\Ndg}$, explaining the asymptote. The resulting localization of the the mode structure to a length of $O(\delta)$, is shown in figure \ref{Asymptotic_mode}, supporting this argument.
